---
title: "R coding Exercise"
output: 
  html_document:
    toc: FALSE
---

```{r read in libraries}
#load dslabs and tidyverse package
library("dslabs")
library("tidyverse")

#look at help file for gapminder data
help(gapminder)

#get an overview of data structure
str(gapminder)

#get a summary of data
summary(gapminder)

#determine the type of object gapminder is
class(gapminder)
```

```{r creating new objects}
#This is filtering the data set into only those who are on the continent of Africa
africadata <- gapminder %>% dplyr::filter(continent == "Africa")

#We are looking at our new object's data structure
str(africadata)

#This is looking at a summary of the data for africadata
summary(africadata)

#This is making subsets of our africadata to only have two variables for each
#infmort only includes infant mortality and life_expectancy from africadata
infmort <- africadata %>% dplyr::select(infant_mortality, life_expectancy)

#pop only includes population size and life_expectancy from africadata 
pop <- africadata %>% dplyr::select(population, life_expectancy)

#We are looking at the data and data structure within both of our new object infmort
#infmort contains the 2907 observations it needs
str(infmort)
summary(infmort)

#We are looking at the data and data structure within both of our new object pop
#pop contains 2907 observations
str(pop)
summary(pop)
```

From the scatter plots, the trend of streaks we are seeing is due to multiple observations from one country by year

```{r scatterplots}
#Scatter plot made using ggplot showing the relationship between infant mortality and life expectancy
ggplot(infmort, aes(x=infant_mortality, y=life_expectancy)) + geom_point()

#Scatter plot was made by using ggplot showing the relationship between population size and life expectancy with a log scale for population
ggplot(pop, aes(x=population, y=life_expectancy)) + geom_point() + scale_x_continuous(trans = "log2")
```

Based on looking at the data structure, there are 51 observations with 9 variables

```{r filter year 2000}
#Filtering to see which years the NA data are and to see there is no NA's for the year 2000
africadata %>% dplyr::filter(is.na(infant_mortality)) %>% dplyr::select(year) %>% count(year)

#Based on the above code, we are filtering africadata to only looking at year 2000 since there is no NA's present
africadata_2000 <- africadata %>% dplyr::filter(year == 2000)

#Looking at the data structure of africadata_2000
str(africadata_2000)

#Looking at a summary of the data of africadata_2000
summary(africadata_2000)
```

According to the scatter plots, infant mortality and life expectancy has a negative correlation while population size and life expectancy has no correlation that we can see. 

```{r scatter plots pt.2}
#Scatter plot for infant mortality and life expectancy based only in the year 2000
ggplot(africadata_2000, aes(x=infant_mortality, y=life_expectancy)) + geom_point()

#Scatter plot for population size and life expectancy based only in the year 2000 with a log scale for population
ggplot(africadata_2000, aes(x=population, y=life_expectancy)) + geom_point() + scale_x_continuous(trans = "log2")
```

Below I will fit two linear models.

```{r linear models}
#A linear model for the africadata_2000 where infant mortality is predicting life expectancy
fit1 = lm(life_expectancy~infant_mortality, data=africadata_2000)

#A linear model for the africadata_2000 where population size is predicting life expectancy
fit2 = lm(life_expectancy~population, data=africadata_2000)

#A summary table displaying information about the linear model between life expectancy and infant mortality
summary(fit1)

#A summary table displaying information about the linear model between life expectancy and population size
summary(fit2)

```

From fit1, we can see that similar to the scatter plot, there is a negative correlation between life expectancy and infant mortality based on the negative estimate for infant_mortality. This estimate is also significant based on the p-value of 2.83e-08 because it is below the common threshold of a p-value of 0.05. This means infant mortality is a significant predictor for life expectancy. For every one unit increase in infant mortality, the life expectancy will decrease 0.189 years.

From fit2, we can see based on the p-value of 0.616, compared to 0.05, that there is no significant linear relationship between the population size and life expectancy. We can see this also based on the estimate for population. It is a value very close to 0. 

#*Additional analysis - Carter Coleman, Group 2*

#To further explore what factors affect life expectancy across African countries in the year 2,000, create a new data set from "africadata_2000" that contains "country", "life expectancy" and "gdp".
```{r}
africadata_2000_life_gdp <- africadata_2000 %>%
  select("country", "life_expectancy", "gdp")
```

#Repeat the last step, except with "Country", "life expectancy", and "Fertility".
```{r}
africadata_2000_life_fert <- africadata_2000 %>%
  select("country", "life_expectancy", "fertility")
```

#Create a plot of life expectancy as the output and country's gdp as the input. Note that whe applied a log(10)scale to gdp to clean up the X-axis a little bit.
```{r}
ggplot(data = africadata_2000_life_gdp) + 
  geom_point(mapping = aes(log10(gdp), life_expectancy))
```

#Repeat the last step with life expectancy and fertility. *Do not apply the log(10) scale to fertility*
```{r}
ggplot(data = africadata_2000_life_fert) + 
  geom_point(mapping = aes((fertility), life_expectancy))
```

#Now that we have some plots, we can see some loose trends: as gdp increses, life expectancy increases (positive relationship); As fertility increases, life expectancy decreases. Since these seem to be some interesting results considering how resource availability affects a country's average life expectancy, fit a linear model to the data sets.

#gdp and life expectancy:
```{r}
fit_gdp_life <- lm(life_expectancy~gdp, data = africadata_2000_life_gdp)
summary(fit_gdp_life)
```
#What we see is that the p-value for the relationship between life expectancy and gdp is <0.05 (p-value = 0.0152) indicating that the relationship is significant. Additionally, the slope (slope = 1.168e-10) of the linear fit shows a positive correltion. However, an R^2 value of 0.096 shows that the linear model does not explain that much of the data. This means while a significant relationship exists between gdp and life expectancy, prediction outside the data set is limited.




#fertility and life expectancy:
```{r}
fit_fert_life <- lm(life_expectancy~fertility, data = africadata_2000_life_fert)
summary(fit_fert_life)
```
#The analysis shows a significant, negative correlation between fertility and life expectancy (p-value = 1.568e-07; slope = -3.8231). In this analysis, R^2 = 0.4211. This means that the linear model explains the data better than that of gdp and life expectancy data. I would argue that because there are so many factors that affect life expectancy, having a R^2 of ~0.4 makes this linear model a decent predictor of the the affect of fertility on life expectancy outside the parameters of this data set.

#Through this exercise, we found three significant relationships that affect life expectancy. Those input variables were infant mortality, fertility, and gdp. Since all those variables do not affect life expectancy in a vacuum, let's look at the relationship of all three concurrently using a multiple regression analysis.

#Use the africa_2000 data set to create a new data set with the variables: country, life expectancy, infant mortality, gdp, and fertility.
```{r}
Multi_africa_2000 <- africadata_2000 %>%
  select("country", "life_expectancy", "fertility", "gdp", "infant_mortality")
```

#*We won't plot this because there are >3 axies being analysed*

#Next, run a multiple regression analysis on the newly created data set with your inputs being: gdp, infant mortality, and fertility; and your output being life expectancy.
```{r}
Multi_fit <- lm(life_expectancy ~ gdp + fertility + infant_mortality, data=Multi_africa_2000)
summary(Multi_fit)
```
#The multiple regression indicated that the model explained 0.4704 of the data. Meaning that, with a p-value = 3.058e-7, the combunation of inputs in the model were a significant predictor of life expectancy. Individually, infant mortality contributed significantly to the model (slope = -1.258e-01, p = 0.0163), gdp (slope = 2.909e-11, p = 0.4641) and fertility (slope = -1.423, p = 0.2074) did not. 


